---
title: "mapping_risk"
author: "Sarah Forrest"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document
---

This code creates a combined map of New York City's death rates per 100,000 and excess risk of dying from dying from COVID-19 by modified zip code tabulation area (modZCTA)

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE, dpi = 300, fig.width = 7)
library(tidyverse)
library(sf)
library(tmap)
library(tmaptools)
library(cowplot)
```

# Read In Data

Read in NYC MODZCTA shapefile:

```{r}
zip_nyc <- st_read("data/mapping_risk/MODZCTA_2010.shp")

str(zip_nyc)

tm_shape(zip_nyc) +
  tm_polygons()
```

Read-in data COVID-19 by NYC MODZCTA:

```{r}
covid <- read_csv("data/mapping_risk/data-by-modzcta.csv")
str(covid)
summary(covid)
```

Join COVID-19 data with NYC MODZCTA shapefile object:

```{r}
covid_zip <- merge(zip_nyc, covid, 
                   by.x = "MODZCTA",
                   by.y = "modified_ZCTA")
```

Generate COVID-19 death rate and relative risk:

```{r}
covid_zip %>% 
  mutate(rate = 100000*(death_count/population),
         expected = population*(35253/8336817),
         RR = death_count/expected) -> covid_zip_rr
```

# COVID-19 Death Rate Map 

```{r}
tm_shape(covid_zip_rr) +
  tm_polygons(col = "rate",
              style = "quantile",
              n = 5,
              palette = "YlOrBr",
              border.col = "black",
              title = "Rate per 100k") +
  tm_legend(legend.position = c(0.12, 0.47),
            legend.text.fontfamily = "Avenir",
            legend.text.size = 1) +
  tm_layout(title = "COVID19 Death Rate \nper 100,000",
            title.size = 1.3,                                                  
            title.position = c(0.10, .85),  
            title.fontfamily = "Avenir",
            frame = FALSE) +
  tm_compass(type = "arrow",
             position = c(0.23, 0.05)) +
  tm_scale_bar(position = c(0.32, 0.04))
```

Save map as an object: 

```{r}
tm_shape(covid_zip_rr) +
  tm_polygons(col = "rate",
              style = "quantile",
              n = 5,
              palette = "YlOrBr",
              border.col = "black",
              title = "Rate per 100k") +
  tm_legend(legend.position = c(0.12, 0.47),
            legend.text.fontfamily = "Avenir",
            legend.text.size = 1) +
  tm_layout(title = "COVID19 Death Rate \nper 100,000",
            title.size = 1.3,                                                  
            title.position = c(0.10, .85),  
            title.fontfamily = "Avenir",
            frame = FALSE) +
  tm_compass(type = "arrow",
             position = c(0.23, 0.05)) +
  tm_scale_bar(position = c(0.32, 0.04)) -> death_rates
```

# COVID-19 Excess Risk Map 

```{r}
tm_shape(covid_zip_rr) +
  tm_polygons(col = "RR",
              n = 5,
              breaks = c(0, 
                         0.50, 
                         0.90, 
                         1.10, 
                         1.50, 
                         2.91),
              palette = "-RdBu",  
              border.col = "black",
              title = "Excess Risk") +
  tm_legend(legend.position = c(0.12, 0.47),
            legend.text.fontfamily = "Avenir",
            legend.text.size = 1) +
  tm_layout(title = "Excess Risk of Dying \nfrom COVID19",
            title.size = 1.3,                                                  
            title.position = c(0.10, .85),  
            title.fontfamily = "Avenir",
            frame = FALSE) +
  tm_credits("Author: Sarah Forrest \nSource: NYC Health \nDate: September 14, 2022",
                        size = .7,
                        position = c(.66,.04))
```

Save map as an object: 

```{r}
tm_shape(covid_zip_rr) +
  tm_polygons(col = "RR",
              n = 5,
              breaks = c(0, 
                         0.50, 
                         0.90, 
                         1.10, 
                         1.50, 
                         2.91),
              palette = "-RdBu",  
              border.col = "black",
              title = "Excess Risk") +
  tm_legend(legend.position = c(0.12, 0.47),
            legend.text.fontfamily = "Avenir",
            legend.text.size = 1) +
  tm_layout(title = "Excess Risk of Dying \nfrom COVID19",
            title.size = 1.3,                                                  
            title.position = c(0.10, .85),  
            title.fontfamily = "Avenir",
            frame = FALSE) +
  tm_credits("Author: Sarah Forrest \nSource: NYC Health \nDate: September 14, 2022",
             size = .7,
             position = c(.66,.04)) -> excess_risk
```

# Combined COVID-19 Death Rate and Excess Risk Map

Turn tmap objects into cowplot objects:

```{r}
tmap_grob(death_rates) -> rates_cow
tmap_grob(excess_risk) -> excess_cow
```

Create a combined figure with cowplot:

```{r}
plot_row <- plot_grid(rates_cow, excess_cow)
```

Add title and save the combined map as an object:

```{r}
title <- ggdraw() + 
  draw_label(
    "New York City COVID-19 Deaths By Zip Code, 2020 - 2021",
    fontfamily = "Avenir",
    x = 0,
    hjust = 0,
    size = 18) +
  theme( plot.margin = margin(0, 0, 0, 25))

plot_grid(title,
          plot_row,
          ncol = 1,
          rel_heights = c(.2,1)) -> combined_cow

combined_cow
```

Save map using `ggsave()`:

```{r}
# ggsave(combined_cow,
#         width = 18, 
#         height = 10,
#         file = "covid_deathrate_excessrisk.jpeg")
```









